--- 
title: "Measuring Precision between Point Clouds based on Different Parameters"
author: "Izaak Cooper"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: ""
---
--- 
title: "Measuring Precision between Point Clouds based on Different Parameters"
author: "Izaak Cooper"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: ""
---

# Abstract

Part of my research for my thesis is analyzing the precision of a point cloud based on the number of ground control points present in a model. The reference model is the model with the most ground control points (this should be the most accurate model). I then compare a test model with less ground control points (less accurate) with the reference model. The way I have been measuring the error between my reference point cloud and my test point cloud has been through a tool that identifies the euclidean distance to the nearest neighbor. Although this tool is useful, I have wondered how much of that Euclidean distance error can be attributed to the X-, Y-, and Z-direction. This is an important subject in my research as the point clouds will be used in bathymetry mapping. In bathymetry mapping, Shifts in the X- and Y- direction would not be as bad as shifts in the Z-direction, which is why I am currently interested in finding how much of the error is in the X-,Y- and Z-direction respectively.


<!--chapter:end:index.Rmd-->

# Introduction {#intro}

You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(methods).

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].

<!--chapter:end:01-intro.Rmd-->

# Literature

Here is a review of existing methods.

<!--chapter:end:02-literature.Rmd-->

# Methods

We describe our methods in this chapter.

<!--chapter:end:03-method.Rmd-->

# Applications

Some _significant_ applications are demonstrated in this chapter.

## Example one

## Example two

<!--chapter:end:04-application.Rmd-->

# Final Words

We have finished a nice book.

<!--chapter:end:05-summary.Rmd-->

`r if (knitr::is_html_output()) '
# References {-}
'`

<!--chapter:end:06-references.Rmd-->

